---
title: "Capstone EDA notebook"
author: "Group 7: Richard Lim, Varun Selvam, Nikita Muddapati, Meenakshi Hariharan"
date: "2025-02-23"
output: 
  html_document:
    theme: flatly
    df_print: paged
    toc: true
    toc_float:
      smooth_scroll: true
---

# EDA Overview

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE) 
```

**Background:**

SCCU(Swire Coca-Cola United States) tries to optimize logistics by transitioning customers selling below a specific annual volume to an Alternate Route to Market (ARTM). There is an annual 400 gallons volume threshold used to distinguish the customers between the direct delivery route and ARTM. However, SCCU is looking for a more cost-efficient strategy to decide new threshold for optimizing logistics which is driving better operational efficiency and more revenues.

**Requirement:**

1.  The analysis will focus on classifying which customers must be included in ARTM or Direct route, and which volume threshold would be optimal to decide for the classification.

2.  The analysis will focus on two key customer segments.

-   1st Group: Local Market Partners that buy fountains only: Customers who buy only fountain drinks and no CO2, cans, or bottles.
-   2nd Group: This group includes all customers, regardless of whether they are local market partners or not, and includes those purchasing CO2, cans, bottles, or fountain drinks.

**Questions:**

-   What factors or characteristics distinguish customers with annual sales exceeding the determined volume threshold from those below this threshold?

-   How can SCCU uses historical sales data, or other Customer Characteristics to predict which ARTM customers have the potential to grow beyond the volume threshold annually?

-   How can these insights be integrated into the routing strategy to support long-term growth while maintaining logistical efficiency?

-   What levers can be employed to accelerate volume and share growth at growth-ready, high-potential customers?

# EDA processes

## 1. Import libraries

```{r}
# import libraries
library(tidyverse)
library(janitor)
library(skimr)
library(psych)
library(glue)
library(here)
library(readxl)
```

## 2. Import Datasets

-   There are 4 datasets used for the analysis, which contains address, customer profile, delivery cost, and transaction history.

```{r}
# import datasets
address_df<- read_csv(here("Dataset",
     "customer_address_and_zip_mapping.csv"))
profile_df <- read_csv(here("Dataset","customer_profile.csv"))
delivery_cost_df <- read_xlsx(here("Dataset","delivery_cost_data.xlsx"))
trans_df <- read_csv(here("Dataset","transactional_data.csv"))
```

## 3. Dataset Profiling & Exploration

### 3-1. Address Dataset Profile

Variables can be described as below.

-   Zip: ZIP code for the location.
-   Full address: Full address information seperated by , including city, state, county, region, and latitude/longitude.
-   Full address is listed in the order of zipcode, city, state full name, state acronym, county, FIPS codes, latitude, longitude

```{r}
sample_n(address_df, 10)
```

### 3-2. Customer Profile Dataset Profile

Variables can be described as below.

-   Customer Number: Unique identifying number of customer
-   Primary Group Number: The group number of which customer mainly belongs to
-   Frequent Order Type: The order type that customer mainly uses
-   First Delivery Date: The date that first delivery was made
-   On Boarding Date: The date that first transaction was made
-   Cold Drink Channel: General channel category for cold drink purchases (e.g., "DINING")
-   Trade Channel: Detailed channel classification (e.g., "OTHER DINING & BEVERAGE")
-   Sub Trade Channel: Sub-classification within the trade channel (e.g., "OTHER DINING")
-   Local Market Partner: Whether customer is local market partner (True or False)
-   CO2 Customer: Whether customer purchases CO2 product or not (True or False)
-   Zip Code: customer address zip code which is connected with Zip variable in `address_df`

```{r}
sample_n(profile_df,10)
```

### 3-3. Delivery Cost Dataset Profile

Variables can be described as below.

-   Cold Drink Channel: The main functional category of commerce
-   Vol Range: The annual volume range of products
-   Applicable to: which category of products that volumes apply to
-   Median Delivery Cost: Median cost of delivery per cost type
-   Cost type: the unit by measuring the cost
    -   Fountain → Measured in gallons (Per Gallon)
    -   Bottles and Cans → Measured in cases (Per Case).

```{r}
sample_n(delivery_cost_df,10)
```

### 3-4. Transaction Dataset Profile

Variables can be described as below.

-   Transaction Date: Date of the transaction (YYYY-MM-DD format).
-   Week: Week number of the year when the transaction occurred.
-   Year: Year of the transaction occurred.
-   Customer Number: Unique identifier for the customer.
-   Order Type: Type of order placed
-   Ordered Cases: The amount of cases that ordered
-   Loaded Cases: The amount of cases that loaded in the truck
-   Delivered Cases: The amount of cases that delivered to the customer
-   Ordered Gallons: The amount of gallons that ordered
-   Loaded Gallons: The amount of gallons that loaded in the truck
-   Delivered Gallons: The amount of gallons that delivered to the customer
    -   **Information 1**: One standard physical case equating to one gallon, allowing for a direct summation of cases and gallons.
    -   **Information 2**: Negative delivered volume must be considered as a return.

```{r}
sample_n(trans_df,10)
```

## 4. Skimming of Dataset

```{r}
skim(address_df)
skim(profile_df)
skim(delivery_cost_df)
skim(trans_df)
```

## 5. Checking NA per variable

```{r}
colSums(is.na(address_df))
colSums(is.na(profile_df))
colSums(is.na(delivery_cost_df))
colSums(is.na(trans_df))
```

-   `PRIMARY_GROUP_NUMBER` has a 18196 missing values, which takes up 60% of `profile_df` dataset.

# EDA questions list

-   How many customers are partnered with Local Market Partners out of the entire customers?

-   How many customers are purchasing C02 products out of entire customers?

-   Which number can we extract out of transaction history?

-   How many customers belongs to the direct route based on the original volume threshold? And how many customers belong to the ARTM based on the original volume threshold?

-   Which customer characteristics have brought more profits from given transaction data?

    -   CO2 vs Non-CO2
    -   Local Market Partners vs Non-Local Market Partners
    -   Cold Drink Channel
    -   Frequent Order Type

-   How many customers belongs to the Local Market Partners that buy fountains only? (Group Segment 1)

-   How many Customers moved above and below the Threshold from 2023 to 2024?

-   What is the Net change in customers moving between threshold categories? (Low Volume, Medium Volume, High Volume)

    -   How many New customers appeared in 2024 compared to 2023?

-   What percentage of customers upgraded or downgraded between categories?

    -   Do customers who move to higher segments tend to have consistent increases in order volume or are they sporadic?
    -   Are there specific patterns in customer order frequency that indicate a transition between volume categories?

-   What are the key patterns in customer order volume reduction from 2023 to 2024.

-   Among customers who reduced their order volume , what is the average percentage drop?

### 1. The summary table of Local Market Partner Customer

```{r}
# the distribution of local market partner customers out of entire customers
table(profile_df$LOCAL_MARKET_PARTNER)
round(prop.table(table(profile_df$LOCAL_MARKET_PARTNER)),2)
```

Approximately, 90% of listed customers belong to the local market partners, which indicates that they are smaller, regionally focused customers who serve their local communities. They tend to show their reliance on local market dynamics and consistent purchasing patterns.

### 2. The summary table of of CO2 customer

```{r}
# the distribution of CO2 customers out of entire customers
table(profile_df$CO2_CUSTOMER)
round(prop.table(table(profile_df$CO2_CUSTOMER)),2)
```

Approximately, 40% of listed customer belongs to the CO2 customer, which represents that they have purchased carbon dioxide materials.

### 3. Total number of transaction

-   Total number of customer
-   Total volume of cases
-   Total volume of gallons
-   Total transaction period

```{r}
trans_df %>%
  summarise(customer_n = n_distinct(CUSTOMER_NUMBER))

trans_df %>%
  summarise(case_volume = sum(ORDERED_CASES),
            gallon_volume = sum(ORDERED_GALLONS),
            total_volume = case_volume + gallon_volume)
```

```{r}
max(as.Date(trans_df$TRANSACTION_DATE, format="%m/%d/%Y"))
min(as.Date(trans_df$TRANSACTION_DATE, format="%m/%d/%Y"))
```

30322 customers have transacted 28,074,470 cases and 10,323,337 gallons (total 38,397,807 units) with SCCU from 1/1/2023 to 12/31/2024. (2 years)

### 4. Transaction history per customer

```{r}
trans_history <-
trans_df %>%
  mutate(TRANSACTION_DATE = as.Date(TRANSACTION_DATE, format="%m/%d/%Y")) %>%
  #mutate(CUSTOMER_NUMBER = as.integer(CUSTOMER_NUMBER)) %>%
  group_by(CUSTOMER_NUMBER) %>%
  summarise(
            FIRST_TRANSACTION_DATE = min(TRANSACTION_DATE),
            LAST_TRANSACTION_DATE = max(TRANSACTION_DATE),
            TRANS_DAYS = LAST_TRANSACTION_DATE - FIRST_TRANSACTION_DATE + 1,
            TRANS_COUNT = n(),
            TRANS_COUNT_2023 = sum((year(TRANSACTION_DATE) == 2023)),
            TRANS_COUNT_2024 = sum((year(TRANSACTION_DATE) == 2024)),
            ANNUAL_VOLUME_CASES_2023 = sum((year(TRANSACTION_DATE) == 2023) * ORDERED_CASES, na.rm = TRUE),
            ANNUAL_VOLUME_GALLON_2023 = sum((year(TRANSACTION_DATE) == 2023) * ORDERED_GALLONS, na.rm = TRUE),
            ANNUAL_VOLUME_CASES_2024 = sum((year(TRANSACTION_DATE) == 2024) * ORDERED_CASES, na.rm = TRUE),
            ANNUAL_VOLUME_GALLON_2024 = sum((year(TRANSACTION_DATE) == 2024) * ORDERED_GALLONS, na.rm = TRUE),
            ANNUAL_VOLUME_2023 = sum((year(TRANSACTION_DATE) == 2023) * (ORDERED_CASES + ORDERED_GALLONS), na.rm = TRUE),
            AVG_ORDER_VOLUME_2023 = ANNUAL_VOLUME_2023 / TRANS_COUNT_2023,
            ANNUAL_VOLUME_2024 = sum((year(TRANSACTION_DATE) == 2024) * (ORDERED_CASES + ORDERED_GALLONS), na.rm = TRUE),
            AVG_ORDER_VOLUME_2024 = ANNUAL_VOLUME_2024 / TRANS_COUNT_2024,
            CHANGED_VOLUME = ANNUAL_VOLUME_2024 - ANNUAL_VOLUME_2023,
            PERCENT_CHANGE = round(CHANGED_VOLUME/ANNUAL_VOLUME_2023,2) * 100,
            THRESHOLD_2023 = ifelse(ANNUAL_VOLUME_2023 >= 400, 'above', 'below'),
            THRESHOLD_2024 = ifelse(ANNUAL_VOLUME_2024 >= 400, 'above', 'below'),
  ) %>%
  ungroup()

trans_history
```

```{r}
colSums(is.na(trans_history))
```

-   calculation of ANNUAL_VOLUME = AVG_ORDER_VOLUME (Order Volume) \* TRANS_COUNT (Frequency) for certain year (2023 vs 2024)

```{r}
# 2023 above vs below threshold
table(trans_history$THRESHOLD_2023)
prop.table(table(trans_history$THRESHOLD_2023))

# 2024 above vs below threshold
table(trans_history$THRESHOLD_2024)
prop.table(table(trans_history$THRESHOLD_2024))
```

-   approximately, 25% of customers are above the original volume threshold (400 annual volume), whereas 75% of customers remain below the threshold in both 2023 and 2024. It appears that the proportion of customer group haven't changed much between 2 years.

```{r}
thres_change_customer <-
trans_history %>%
  filter(THRESHOLD_2023 != THRESHOLD_2024)

thres_change_customer
```

```{r}
table(thres_change_customer$THRESHOLD_2023, thres_change_customer$THRESHOLD_2024)
round(prop.table(table(thres_change_customer$THRESHOLD_2023, thres_change_customer$THRESHOLD_2024)),2)
```

However, when we get into the depth, 2,378 (8%) customers experienced a change in volume based on the original volume threshold from 2023 to 2024 out of 30,322 total customers. Among them, 1,250 customers (around 4%) exceeded the threshold in 2024 from below threshold status, whereas 1,128 (around 4%) customers drops below the threshold.

## 5. Volume changes comparison

### 5-1. Changed volume statistics

```{r}
# total customer growth statistics
trans_history %>%
  summarise(AVG_CHANGE_VOL = mean(CHANGED_VOLUME),
            MED_CHANGE_VOL = median(CHANGED_VOLUME),
            MIN_CHANGE_VOL = min(CHANGED_VOLUME),
            MAX_CHANGE_VOL = max(CHANGED_VOLUME))

# below in both year growth statistics

trans_history %>%
  filter(THRESHOLD_2023 == 'below' & THRESHOLD_2024 == 'below') %>%
  summarise(AVG_CHANGE_VOL = mean(CHANGED_VOLUME),
            MED_CHANGE_VOL = median(CHANGED_VOLUME),
            MIN_CHANGE_VOL = min(CHANGED_VOLUME),
            MAX_CHANGE_VOL = max(CHANGED_VOLUME))

# above in both year growth statistics

trans_history %>%
  filter(THRESHOLD_2023 == 'above' & THRESHOLD_2024 == 'above') %>%
  summarise(AVG_CHANGE_VOL = mean(CHANGED_VOLUME),
            MED_CHANGE_VOL = median(CHANGED_VOLUME),
            MIN_CHANGE_VOL = min(CHANGED_VOLUME),
            MAX_CHANGE_VOL = max(CHANGED_VOLUME))

# potential growth customer statistics
trans_history %>%
  filter(THRESHOLD_2023 == 'below' & THRESHOLD_2024 == 'above') %>%
  summarise(AVG_CHANGE_VOL = mean(CHANGED_VOLUME),
            MED_CHANGE_VOL = median(CHANGED_VOLUME),
            MIN_CHANGE_VOL = min(CHANGED_VOLUME),
            MAX_CHANGE_VOL = max(CHANGED_VOLUME))
```

### 5-2. Changes in volume percent distribution

```{r}
# total customer
trans_history %>%
  ggplot() +
  geom_boxplot(aes(x = PERCENT_CHANGE)) +
  theme_minimal()

# both below customer
trans_history %>%
  filter(THRESHOLD_2023 == 'below' & THRESHOLD_2024 == 'below') %>%
  ggplot() +
  geom_boxplot(aes(x = PERCENT_CHANGE), na.rm = TRUE) +
  theme_minimal()

# both above customer
trans_history %>%
  filter(THRESHOLD_2023 == 'below' & THRESHOLD_2024 == 'below') %>%
  ggplot() +
  geom_boxplot(aes(x = PERCENT_CHANGE), na.rm = TRUE) +
  theme_minimal()

# potential growth customer
trans_history %>%
  filter(THRESHOLD_2023 == 'below' & THRESHOLD_2024 == 'above') %>%
  ggplot() +
  geom_boxplot(aes(x = PERCENT_CHANGE)) +
  theme_minimal()
```

### 6. Combining the Dataset (Data Modeling)

In order to take in-depth analysis per each of customer's attributes, we've combined the customer profile `profile_df` data with `trans_history` , joined by `CUSTOMER_NUMBER` variable.

```{r}
trans_profile_df <- left_join(trans_history, profile_df, by = 'CUSTOMER_NUMBER')
sample_n(trans_profile_df,10)
```

## Variable comparison analysis

### 7-1. Local Market Partner Comparison

```{r}
volume_2023 <- sum(trans_profile_df$ANNUAL_VOLUME_2023, na.rm = TRUE)
volume_2024 <- sum(trans_profile_df$ANNUAL_VOLUME_2024, na.rm = TRUE)

trans_profile_df %>%
  group_by(LOCAL_MARKET_PARTNER) %>%
  summarise(TOTAL_VOL_2023 = sum(ANNUAL_VOLUME_2023),
            TOTAL_VOL_2024 = sum(ANNUAL_VOLUME_2024),
            PERCENT_2023 = (TOTAL_VOL_2023 / volume_2023) * 100,
            PERCENT_2024 = (TOTAL_VOL_2024 / volume_2024) * 100,
            AVG_VOL_2023 = mean(ANNUAL_VOLUME_2023),
            AVG_VOL_2024 = mean(ANNUAL_VOLUME_2024),
            MED_VOL_2023 = median(ANNUAL_VOLUME_2023),
            MED_VOL_2024 = median(ANNUAL_VOLUME_2024),
            COUNT_2023 = sum(TRANS_COUNT_2023),
            COUNT_2024 = sum(TRANS_COUNT_2024),
            ABOVE_THRES_2023 = sum(THRESHOLD_2023 == 'above'),
            ABOVE_THRES_2024 = sum(THRESHOLD_2024 == 'above')
  )
```

### 7-2. C02 customer Comparison

```{r}
trans_profile_df %>%
  group_by(CO2_CUSTOMER) %>%
  summarise(TOTAL_VOL_2023 = sum(ANNUAL_VOLUME_2023),
            TOTAL_VOL_2024 = sum(ANNUAL_VOLUME_2024),
            PERCENT_2023 = (TOTAL_VOL_2023 / volume_2023) * 100,
            PERCENT_2024 = (TOTAL_VOL_2024 / volume_2024) * 100,
            AVG_VOL_2023 = mean(ANNUAL_VOLUME_2023),
            AVG_VOL_2024 = mean(ANNUAL_VOLUME_2024),
            MED_VOL_2023 = median(ANNUAL_VOLUME_2023),
            MED_VOL_2024 = median(ANNUAL_VOLUME_2024),
            COUNT_2023 = sum(TRANS_COUNT_2023),
            COUNT_2024 = sum(TRANS_COUNT_2024),
            ABOVE_THRES_2023 = sum(THRESHOLD_2023 == 'above'),
            ABOVE_THRES_2024 = sum(THRESHOLD_2024 == 'above')
  )
```

### 7-3. Frequent order type Comparison

```{r}
trans_profile_df %>%
  group_by(FREQUENT_ORDER_TYPE) %>%
  summarise(TOTAL_VOL_2023 = sum(ANNUAL_VOLUME_2023),
            TOTAL_VOL_2024 = sum(ANNUAL_VOLUME_2024),
            PERCENT_2023 = (TOTAL_VOL_2023 / volume_2023) * 100,
            PERCENT_2024 = (TOTAL_VOL_2024 / volume_2024) * 100,
            AVG_VOL_2023 = mean(ANNUAL_VOLUME_2023),
            AVG_VOL_2024 = mean(ANNUAL_VOLUME_2024),
            MED_VOL_2023 = median(ANNUAL_VOLUME_2023),
            MED_VOL_2024 = median(ANNUAL_VOLUME_2024),
            COUNT_2023 = sum(TRANS_COUNT_2023),
            COUNT_2024 = sum(TRANS_COUNT_2024),
            ABOVE_THRES_2023 = sum(THRESHOLD_2023 == 'above'),
            ABOVE_THRES_2024 = sum(THRESHOLD_2024 == 'above')
  )
```

### 7-4. Cold Drink Channel Comparison

```{r}
trans_profile_df %>%
  group_by(COLD_DRINK_CHANNEL) %>%
  summarise(TOTAL_VOL_2023 = sum(ANNUAL_VOLUME_2023),
            TOTAL_VOL_2024 = sum(ANNUAL_VOLUME_2024),
            PERCENT_2023 = (TOTAL_VOL_2023 / volume_2023) * 100,
            PERCENT_2024 = (TOTAL_VOL_2024 / volume_2024) * 100,
            AVG_VOL_2023 = mean(ANNUAL_VOLUME_2023),
            AVG_VOL_2024 = mean(ANNUAL_VOLUME_2024),
            MED_VOL_2023 = median(ANNUAL_VOLUME_2023),
            MED_VOL_2024 = median(ANNUAL_VOLUME_2024),
            COUNT_2023 = sum(TRANS_COUNT_2023),
            COUNT_2024 = sum(TRANS_COUNT_2024),
            ABOVE_THRES_2023 = sum(THRESHOLD_2023 == 'above'),
            ABOVE_THRES_2024 = sum(THRESHOLD_2024 == 'above')
  )
```

### 8. Group Segment #1

```{r}
# Group 1: Local Market Partners that buy fountains only
group1_df <-
trans_profile_df %>%
  filter(!CO2_CUSTOMER 
         & LOCAL_MARKET_PARTNER 
         & ANNUAL_VOLUME_CASES_2023 == 0
         & ANNUAL_VOLUME_CASES_2024 == 0)

group1_df %>%
  summarise(TOTAL_VOLUME_2023 = sum(ANNUAL_VOLUME_GALLON_2023),
            TOTAL_VOLUME_2024 = sum(ANNUAL_VOLUME_GALLON_2024),
            ABOVE_THRES_2023 = sum(THRESHOLD_2023 == 'above'),
            ABOVE_THRES_2024 = sum(THRESHOLD_2024 == 'above'))
```

### 9. Threshold Comparison for 2023 and 2024

```{r}
# Define threshold (400 gallons)
threshold <- 400  

# Filter data for 2023 and 2024 only
transaction_filtered <- trans_df %>%
  filter(YEAR %in% c(2023, 2024)) 

# Summarize transactions per customer per year
customer_summary <- transaction_filtered %>%
  group_by(CUSTOMER_NUMBER, YEAR) %>%
  summarise(
    Total_Ordered_Cases = sum(ORDERED_CASES, na.rm = TRUE),
    Total_Ordered_Gallons = sum(ORDERED_GALLONS, na.rm = TRUE),
    Order_Frequency = n(),
    .groups = "drop"
  ) %>%
  # Add Total Volume Calculation
  mutate(
    Total_Volume = Total_Ordered_Cases + Total_Ordered_Gallons,
    Customer_Category = ifelse(Total_Ordered_Gallons >= threshold, "Above Threshold", "Below Threshold")
  ) %>%
  
  # Volume Segmentation
  mutate(
    Volume_Segment = case_when(
      Total_Volume >= 1000 ~ "High Volume",
      Total_Volume >= 500 ~ "Medium Volume",
      TRUE ~ "Low Volume"
    )
  ) %>%
  #  missing values
  mutate(
    Customer_Category = replace_na(Customer_Category, "Unknown"),
    Volume_Segment = replace_na(Volume_Segment, "Unknown")
  )

# Customers who changed from 2023-2024
threshold_change_customers <- customer_summary %>%
  select(CUSTOMER_NUMBER, YEAR, Customer_Category) %>%
  pivot_wider(names_from = YEAR, values_from = Customer_Category, values_fill = list(Customer_Category = "No Purchase")) %>%
  rename(Threshold_2023 = `2023`, Threshold_2024 = `2024`) %>%
  filter(Threshold_2023 != Threshold_2024)
         
threshold_transition_summary <- threshold_change_customers %>%
  group_by(Threshold_2023, Threshold_2024) %>%
  summarise(Customers_Transitioned = n(), .groups = "drop")

# Calculate net change in threshold categories
net_change_summary <- threshold_transition_summary %>%
  mutate(Change = case_when(
    Threshold_2023 == "Below Threshold" & Threshold_2024 == "Above Threshold" ~ Customers_Transitioned,
    Threshold_2023 == "Above Threshold" & Threshold_2024 == "Below Threshold" ~ -Customers_Transitioned,
    TRUE ~ 0
  )) %>%
  summarise(Net_Change = sum(Change))

# Track Customers Who Changed Volume Segments (Low/Medium/High)
volume_change_customers <- customer_summary %>%
  select(CUSTOMER_NUMBER, YEAR, Volume_Segment, Total_Volume, Order_Frequency) %>%
  pivot_wider(names_from = YEAR, values_from = c(Volume_Segment, Total_Volume, Order_Frequency),
              values_fill = list(Volume_Segment = "No Purchase", Total_Volume = 0, Order_Frequency = 0)) %>%
  rename(Volume_2023 = Volume_Segment_2023, Volume_2024 = Volume_Segment_2024,
         Volume_Ordered_2023 = Total_Volume_2023, Volume_Ordered_2024 = Total_Volume_2024,
         Order_Frequency_2023 = Order_Frequency_2023, Order_Frequency_2024 = Order_Frequency_2024)

# Identify customers with consistent or sporadic increases
volume_growth_analysis <- volume_change_customers %>%
  filter(Volume_2023 != "No Purchase" & Volume_2024 != "No Purchase" & Volume_2023 != Volume_2024) %>%
  mutate(Volume_Growth_Trend = case_when(
    Volume_Ordered_2024 > Volume_Ordered_2023 ~ "Consistent Growth",
    Volume_Ordered_2024 < Volume_Ordered_2023 ~ "Fluctuating",
    TRUE ~ "Stable"
  ))

# Identify patterns in customer order frequency changes
order_frequency_analysis <- volume_change_customers %>%
  mutate(Frequency_Change = Order_Frequency_2024 - Order_Frequency_2023,
         Frequency_Pattern = case_when(
           Frequency_Change > 0 ~ "Increasing Frequency",
           Frequency_Change < 0 ~ "Decreasing Frequency",
           TRUE ~ "Stable Frequency"
         ))

# Calculate average percentage drop for customers who reduced order volume
order_volume_drop_analysis <- volume_change_customers %>%
  filter(Volume_Ordered_2023 > 0 & Volume_Ordered_2024 < Volume_Ordered_2023) %>%
  mutate(Percentage_Drop = case_when(
    Volume_Ordered_2023 > 0 ~ ((Volume_Ordered_2023 - Volume_Ordered_2024) / Volume_Ordered_2023) * 100,
    TRUE ~ NA_real_  # Avoid division by zero
  ))

# Print summaries
print(order_frequency_analysis)
print(order_volume_drop_analysis)
```

### 10. Visualization for Order Volume

```{r}
# Visualization for Order Volume Drop Distribution
ggplot(order_volume_drop_analysis, aes(x = Percentage_Drop)) +
  geom_histogram(binwidth = 5, fill = "purple", alpha = 0.7, color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Order Volume Drop (2023 → 2024)",
       x = "Percentage Drop in Order Volume",
       y = "Number of Customers")
```

### 11. Visualization for Customer Order Frequency Changes

```{r}

ggplot(order_frequency_analysis, aes(x = Frequency_Pattern, fill = Frequency_Pattern)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Customer Order Frequency Changes (2023 → 2024)",
       x = "Order Frequency Pattern",
       y = "Number of Customers",
       fill = "Frequency Change") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### 12. Visualization for Volume Growth Trends

```{r}
ggplot(volume_growth_analysis, aes(x = Volume_Growth_Trend, fill = Volume_Growth_Trend)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Customer Volume Growth Trends (2023 → 2024)",
       x = "Growth Trend",
       y = "Number of Customers",
       fill = "Trend Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# EDA Insights (Summary)

Out of EDA, we could find out below insights

-   There are 30,322 unique customers from 2023/01/01 to 2024/12/31. (2 years) out of transaction history data.

-   Even if approximately 90% of customers belongs to the Local Market Customers, their total volume of transaction takes up 72% of entire transaction volumes.

    -   There is an 2% point increase of proportion in 2024 for Local Market Customers compared to 2023, which represents local retails growth potential.
    -   Local market customers are likely to order +4 more frequencies with +4 less volume compared to non-local market customer.

-   Even though there is not much change of ordering pattern between CO2 customer and Non CO2 customer in 2023 and 2024, median volume per order has increased by over 10% in 2024 compared to 2023 for C02 customer.

-   SALES REP (sales representatives) remains in 75% of order type for 2 years transactions, followed by OTHERS, and MYCOKE360 (Digital Ordering Platform), which indicates that personal interaction is still significant to maintain the sales.

    -   However, EDI ordering volume increase over 2 times more, and MYCOKE360 volumes increase by 1.5 times from 2023 to 2024.

-   In terms of order volume percentage per year, Goods channel increase by 2% points from 2023 to 2024.

-   BULK TRADES and DINING takes over 50% of entire transaction volume in both 2023 and 2024.

-  14,742 customers experienced a decline in order volume, including some high-volume customers moving to medium or low volume.
Growth segment: Certain customers moved from low to high volume, indicating rising demand and potential need for priority servicing.

- Some customers crossed above or below the 400-gallon threshold, affecting route efficiency and delivery planning.
Net Impact: Helps assess whether SCCU should expand direct delivery routes or refine ARTM logistics.

- Increased order frequency suggests growth potential, while decreased frequency may signal churn risk.

# Contribution

- Richard Lim: Structuring and organizing the EDA notebook
- Varun Selvam: Yaml file formatting and data validation
- Nikita Muddapati: Delivery cost calculation and additional EDA questions
- Meenakshi Hariharan: Implementing threshold, volume analysis and key patterns in customer order, volume reduction from 2023 to 2024
